---
title: "lab_02"
author: "derek willis"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1. 26 departments offer these courses, and the Atmospheric and Oceanic Science department offers the most with 68.**

```{r}
umd_courses |>
  filter(str_detect(title, "Climate")& seats > 0) |>
  group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes)) 
  
  
```

**Q2.** Of the departments listed in A2, which one is the most interesting to you, and why? **A2. I think which one is the most interesting to me is Economics, because when I think of economic policy, I feel like they actively seek to destroy the environment, especially the atmosphere and the ocean.**

**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result?

**A3. Business Management and Economics have the most courses in this term. 139 departments in that term have at least one syllabus posted. The top five departments are different in each result because Business Management has 186 classes, Economics has 129 courses, Biological Sciences has 109 classes, Information Sciences has 104, and School of Music has 99.**

```{r}
umd_courses |>
  filter(syllabus_count >= 1) |>
  group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))
  
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time. **A4. The pattern of spending over time isn't linear in term of increasing, but the higher values definitely are present in 2022, meaning there is an increase over time.**

```{r}
maryland_expenses |>
  mutate(expenditure_date = mdy(expenditure_date)) |>
  mutate(month = floor_date(expenditure_date, "month")) |>
#filter(commitee_name == 'Moore Wes For Maryland') |>
group_by(month) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5 I had absolutely no idea how to do this, and tried to figure it out and deleted everything. Like oh my.**

```{r}
maryland_expenses |>
  
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6. The state that received the least amount of aid from Maryland in the expenses with states was West Virginia.**

```{r}
maryland_expenses_with_state |>
  group_by(state) |>
  summarize(total = sum(amount)) |>
  arrange((total))
```
