---
title: "Reverse Engineering Project Notebook"
output: html_notebook
---

Deliverable 5: Final Notebook/Memo <https://docs.google.com/document/d/1197C83YSa2MKSKzEyou_yHwQQzzwBio_74Ynt1MvMDQ/edit?usp=sharing>

Load libraries.

```{r}
options(scipen=999)

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
```

Load data. Change amount column from character to a number.

```{r}

#change Amount from <chr> to <dbl> by getting rid of special characters

sportsdonation <- read_csv("data/sports-political-donations.csv") |>
  clean_names() |>
  mutate(amount = str_replace_all(amount,'\\$','')) |>
  mutate(amount = str_replace_all(amount, '\\,','')) |>
  mutate(amount = as.numeric(amount))
  
sportsdonation
```

1.  "Of the 160 owners and commissioners in ESPN's database, 74 contributed a majority of their publicly searchable funds to Republican campaigns, with 48 donating a majority to Democrats and the rest sending most of their money to causes that aren't fully aligned with either party."

```{r}

#number of owners/commissioners —— 158 rows, have to figure out way to get both owners and commisioners for the "more than 160 owners and commisioners" This might be because
sportsdonation |>
  filter(str_detect(party, 'Republican')) |>
  group_by(owner) |> 
  summarise(total = n())

```

2.  "American professional sports owners have contributed **nearly \$47 million in federal elections** since 2015, according to research by ESPN in partnership with FiveThirtyEight, including **\$10 million to Republican causes** and **\$1.9 million to Democratic causes** so far in the 2020 election cycle."

```{r}

#sum total donations: $46,978,697 ~ nearly $47 million.

sportsdonation |>
  summarise(amount) |>
  sum()

#find total Democrat donations specific to 2020 election cycle: $1,877,133 ~ $1.9 million.

sportsdonation |>
  filter(str_detect(party, "Democrat")) |>
  filter(election_year== "2020") |>
  summarise(amount) |>
  sum()

#find total Republican donations specific to 2020 election cycle: $10,111,284 ~ $10 million.

sportsdonation |>
  filter(str_detect(party, "Republican")) |>
  filter(election_year == "2020") |>
  summarise(amount) |>
  sum()

```

3.  **Which team owners gave the most?**

    Donations from team owners, by cyle and party donated to: (this was in a table in the original story)

```{r}

# Who provided the highest number of donations? Charles Johnson.
sportsdonation |>
  group_by(owner) |>
  summarise(total_donations = n()) |>
  arrange(desc(total_donations))

# Which league gave the most individual donations? MLB
sportsdonation |>
  group_by(league) |>
  summarise(total_donation = n()) |>
  arrange(desc(total_donation))

```

4.  Overall, **51.7 percent of all contributions by WNBA owners went to Democrats**, with **42.3 percent going to Republicans.** And of the **\$1.3 million contributed by WNBA owners to Republicans, 65.5 percent comes from Kelly Loeffler,** the Republican from Georgia who is co-owner of the Atlanta Dream.

```{r}


sportsdonation |>
  filter(league == 'WNBA') |>
  filter(owner == 'Kelly Loeffler') |>
  summarise(wnba_amt = sum(amount)) |>
  arrange(desc(wnba_amt))

#Kelly Loeffler: $896,150

sportsdonation |>
  filter(league == 'WNBA') |>
  filter(str_detect(party, 'Republican')) |>
  summarise(sum(amount))

#when adding up Republican WNBA spending, the total is $1,055,552 — this is not $1.3m, have to figure out why. We are not sure what the discrepancy is here, but we think it might be because there is not data for "Bipartisan, but mostly Democrat" included, which makes the total a little less than what FiveThirtyEight calculated. 

#FiveThirtyEight had a searchable database, and we found that some of the donations that were there were not present in our data, particularly  Kelly Loeffler's bipartisan donations. Even when we tried to redownload the data, we got the same results, so this seems to be something with the data.

100*(896150/1055552)

100*(896150/1300000)

sportsdonation |>
  filter(league == 'WNBA') |>
  group_by(party) |>
  summarise(total = sum(amount))
# there seems to be no WNBA donations for "Bipartisan, but mostly Democrat." 

sportsdonation |>
  filter(league == 'WNBA') |>
  group_by(owner) |>
  summarise(total = sum(amount))

```

5.  "There are 40 owners who have donated at least \$100,000 to Republican causes over the six-year period, compared to 23 owners topping \$100,000 to Democrats, helping lead to these splits."

```{r}

# There are 37 rows of owners that donated more than $100,000 to Republican causes.
sportsdonation |>
  filter(str_detect(party, 'Republican')) |>
  group_by(owner) |>
  summarise(total = sum(amount)) |>
  filter(total >= 100000)

# 22 rows of owners that donated more than $100,000 to Democrat causes.
sportsdonation |>
  filter(str_detect(party, 'Democrat')) |>
  group_by(owner) |>
  summarise(total = sum(amount)) |>
  filter(total >= 100000)

#These discrepancies are a little odd, as they are just slightly off from what FiveThirtyEight reported. We think that this could be because of the same issue with the WNBA calculations due to the data itself.


# Which year had the most donations? 2018
sportsdonation |>
  group_by(election_year) |>
  summarise(count_per_year = n()) |>
  arrange(desc(count_per_year))

```

6.  "A deep search in the Federal Election Commission database of campaign finances for principal owners, controlling owners, co-owners and commissioners from the NBA, NFL, NHL, WNBA, MLB and NASCAR reveals that this deep-pocketed group has sent **\$34.2 million (72.9 percent) to Republican campaigns or super PACs purely supporting Republican** causes, compared to \$10.1 million (21.5 percent) to Democrats over the past three elections. Less than 6 percent of contributions went to bipartisan or unaffiliated recipients."

```{r}

# Money sent to purely Republican causes: 
# sum: 34,246,016 ~ $34.2 million, the total donation amount was $46,978,697
# 34,246,016/46,978,697 = 0.728969 = 72.9%
sportsdonation |>
  filter(party == 'Republican') |>
  summarise(sum(amount))

100*(34246016/46978697)

# Money sent to purely Democrat causes:
# sum: $10,113,639 ~ $10.1 million.
# $10,113,639/46,978,697 = 0.2152814 = 21.5%

sportsdonation |>
  filter(party == 'Democrat') |>
  summarise(sum(amount))

100*(10113639/46978697)
```

```{r}

#data visualization

league_donation <- sportsdonation |>
  group_by(league) |>
  summarise(total_donation = n()) |>
  arrange(desc(total_donation))

league_donation |>
  ggplot() +
  geom_bar(aes(x=league, weight=total_donation)) +
  theme(
    axis.text.x = element_text(angle = 45,  hjust=1)
  ) +
   labs(
    title="Political Donation by Sports League",
    x = "League",
    y = "Total Number of Donation",
    caption = "source: FiveThirtyEight and ESPN"
  ) 



```
