---
title: "lab_01"
author: "Derek Willis"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

**Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?
**A1** The one that is most surprising to me is that Air Science has the sixth most mean seats in the school. I've never heard of that department before and wasn't even sure that it offerred a major.

```{r}
umd_courses |>
group_by(department) |>
summarise (
count_classes = n(),
total_seats = sum(seats),
mean_seats = mean(seats), 
 ) |>
+ arrange(desc(mean_seats))
```

**Q2** Find out the greatest number of times any individual class has been offered in our dataset. Your code must show the greatest number of classes first. 
**A2**

```{r}
 umd_courses |>
 group_by(id) |>
 summarise(
  count_classes = n(),
) |>
arrange(desc(count_classes))
```

**Q3** Count the number of courses for each instructor and department. Which instructor has taught the most? **A3** This data says that Keith Yearwood has taught the most classes.

```{r} 
umd_courses |>
group_by(instructors, department) |>
summarise(
count_classes = n(),
) |>
arrange(desc(count_classes))
```

**Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see. How confident are you that the answer you got is the right one, and why? **A4** What would have made it easier to answer q3 is if all the N/A titles in the professors data set could've been taken out of the results. I'm pretty confident the answer that I got is the right one, because I weeded through all of the data and found the first name that wasn't N/A from my results. Although now thinking about it, the N/A could've been a professor.

**Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting? **A5** All of my questions that I asked on the pre lab require extra data that is not included in this data set. I would need enrollment numbers in the university, and what percent of thos seats are filled each semester on average for each department. Although I guess it would look something like this for the first question.

```{r}
umd_courses |>
group_by(term, department) |>
summarise (
count_classes = n(),
total_seats = sum(seats),
mean_seats = mean(seats), 
 ) |>
arrange(desc(mean_seats))
```
