---
title: "week7_recap"
author: "Daniel Trielli"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load Data

Run the codeblock below to load the data.

```{r}
earthquakes <- read_csv('https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv')

#Setting time column as datetime
earthquakes <- earthquakes |> mutate(time = as_datetime(time))
```

#### Answer the questions below

Most questions have a code block and a space for an answer below. Write the code you think is necessary and, in the answer space, write out what you did and what was the result.

------------------------------------------------------------------------

#### **Q1** Do you notice anything different with this read_csv()? Particularly in the file name?

**A1: In the file name there are no underscores and it is just one name, and in the csv file there is a link to a website instead of it just being a csv file. There are also a lot of columns with mostly numbers.**

------------------------------------------------------------------------

#### **Q2** How many records there are there in this dataset? What do they mean and what useful information we can gather from it, looking at the columns?

**A2: There are 9,776 records in this dataset. They are each records of different earthquakes that occurred around the world on a daily basis that updates every minute. Useful information we can gather are the place the earthquake happened, the magnitude of the earthquake and the depth.**

------------------------------------------------------------------------

#### **Q3** How do I reorganize this list to see the ones that are the deepest first? What is the depth that shows up for the deepest one, and its magnitude?

```{r}
earthquakes |> 
  arrange(desc(depth)) 
  
```

**A3: 669.9820 is the depth that shows up for the deepest one, and it has a magnitude of 4.20.**

------------------------------------------------------------------------

#### **Q4** I just want to see the earthquakes with a magnitude larger than 6. How do I do that? And how many are there that fit this criteria?

```{r}
earthquakes |>
  filter( mag > 6)
           
```

**A4: There are 16.**

------------------------------------------------------------------------

#### **Q5** What about if I want to see earthquakes that have both a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [both]{.underline} these criteria?

```{r}
earthquakes |>
  filter( mag > 6 & depth < 20)

```

**A5: Filter by and. There are only five earthquakes in the data set that match this criteria.**

------------------------------------------------------------------------

#### **Q6** What about if I want to see earthquakes that either have a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [either]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag >= 6 | depth <= 20)



```

**A6: Filter with or \|, there are 7,423 earthquakes that match that description.**

------------------------------------------------------------------------

#### **Q7** I'm interested in finding earthquakes that took place in Alaska. Which column should I look at? How do I use it to find all the earthquakes in Alaska? How many earthquakes took place there?

```{r}
earthquakes |> filter(str_detect(place, "Alaska"))
  
```

**A7: You would like to look at the place column. Then you str detect it. There are 3, 423 earthquakes that took place in Alaska.**

------------------------------------------------------------------------

#### **Q8** I notice that there is a column called 'type', that seems to have different kinds of tremors. What is the best way to find what are all the possible types of tremors, and counting how many of each there are in this data set? What are the first two most common types of tremors in this data set?

```{r}
earthquakes |> group_by(type) |> summarise(count_type =n())
```

**A8: First most common are earthquakes, second most common are quary blasts. The best way is to group by and then summarise everything.**

------------------------------------------------------------------------

#### **Q9** How do I calculate the average depth of each type of tremor? What is the average depth of the earthquake type in this data set? Is there anything that seems unusual or surprising in this finding?

```{r}
earthquakes |>
  group_by(type) |>
  summarise(average_depth = mean(depth))
```

**A9: You group by and then summarise to find the mean(depth). The thing that seems unusual is that there is negative depth for some of these earthquakes, which means the ground rose up. The average depth of the earthquake type is 25.28 meters.**

------------------------------------------------------------------------

#### **Q10** I'm interested, in the future, to see the hours in which earthquakes happen. How can I extract the hour from the time column?

```{r}
earthquakes <- earthquakes |>
  mutate(hour = hour(time))
```

**A10: You mutate the time column into a new column, using the time function.**

------------------------------------------------------------------------

#### **Q11** I want to make a note of all the records in this data set that I consider serious. For my purposes, I'm thinking that all tremors that have a magnitude that is larger than 3 are serious. How do I automatically create a new column making that distinction?

```{r}
earthquakes <- earthquakes |>
  mutate(serious = case_when(
    mag > 3 ~ 'Serious',
    .default = 'Not serious'
  ))
```

**A11: You create a new column by mutating the data and then using a case_when to make the distinction between the two.**

------------------------------------------------------------------------

#### **Q12** I have no idea how earthquakes work and I'm interested in seeing if there is a particular time of day in which serious earthquakes happen. How can I see that condensed in a table with all the hours in a day and all the serious earthquakes in each hour? What is the hour with fewer serious earthquakes and the one with the most serious earthquakes?

```{r}
earthquakes <- earthquakes |>
  filter(serious == 'Serious' & type =='earthquake') |>
  group_by(hour) |>
  summarise(count = n()) |>
  arrange(desc(count))

```

**A12**: You can filter by the serious column you just created, and then group by the hour and summarise by n. 2 is the hour with the most serious earthquakes and 12 is the hour with the least serious earthquakes.

------------------------------------------------------------------------

#### EXTRA Q What other questions are you interested in and how would you ask them in R?

```{r}

```

```{r}

```

```{r}

```
